---
title: "Data 612 Project 1 - Global Baseline Predictors and RMSE"
author: "Vinayak Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: tango
    theme: united
  word_document: default
  pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```


```{r message=FALSE, warning=FALSE}
if (!require("knitr")) install.packages("knitr")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("dplyr")) install.packages("dplyr")
```

<p>&nbsp;</p>

## Introduction

**Briefly describe the recommender system that you’re going to build out from a business perspective, e.g. “This system recommends data science books to readers.”**

<span style="font-size:16px;"> This system recommends members watched movies to other family members </span>

**Find a dataset, or build out your own toy dataset. As a minimum requirement for complexity, please include numeric ratings for at least five users, across at least five items, with some missing data.**

<span style="font-size:16px;"> For dataset, I created a survey in family, and I asked them to give ratings to movie which was picked by family members.

There are 6 members and 8 movies. </span>

## Data 

**Load your data into (for example) an R or pandas dataframe, a Python dictionary or list of lists, (or another data structure of your choosing).** 

```{r echo=TRUE}
#Load results
results <- read.csv("https://raw.githubusercontent.com/Vinayak234/Data612/master/movie_ratings.csv", header = TRUE)

results %>% kable()%>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```

**The dimensions of the results dataframe are (`r dim(results)`)**

```{r}
### Structure
str(results)
### Summary
summary(results)
```

## Transform Data

I use `pivot_wider` to convert the above format into a table with 8 rows and 6 columns.

```{r}
#### use pivot_wider to make a user matrix
results %>% pivot_wider(names_from = Movie_Name, values_from = Ratings) -> UI
UI  %>% column_to_rownames("User_Name") -> UI
UI
```


## Split training and test datasets.

I will make a matrix of ones and zeros which will facilitate extracting the desired elements from the overall matrix.   

#### Test dataset:

```{r}
test_rows <- c(1, 2, 3, 4, 5, 6, 5, 4)
test_cols <- c(3, 4, 6, 8, 5, 1, 2, 7)
test_extractor <- cbind(test_rows, test_cols)

UI_train <- UI
UI_train[test_extractor] <- NA
UI_train <- as.matrix(UI_train)

UI_test <- UI
UI_test[test_extractor] <- 0
UI_test[UI_test > 0] <- NA
UI_test[test_extractor] <- UI[test_extractor]
UI_test <- as.matrix(UI_test)

### display UI_train
UI_train %>% kable(caption = "TRAINING MATRIX")%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

### display UI_test
UI_test %>% kable(caption = "TEST MATRIX")%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```


## Using your training data, calculate the raw average (mean) rating for every user-item combination.

```{r}
mean_value <- raw_ave <- mean(UI_train, na.rm = TRUE)
mean_value

### make a matrix with same rownames and colnames as UI_train, but replace the values
mean_rating <- UI_train
mean_rating[T] <- mean_value
mean_rating  %>% kable(caption = "MEAN-RATING MATRIX")%>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```



## Calculate the RMSE for raw average for both your training data and your test data.

```{r}
### Training RMSE 
train_RMSE_raw <- sqrt( mean ( (UI_train - mean_rating)^2, na.rm=T ) )
train_RMSE_raw

### Test RMSE 
test_RMSE_raw <- sqrt( mean ( (UI_test - mean_rating)^2, na.rm=T ) )
test_RMSE_raw

```



### Using your training data, calculate the bias for each user and each item.

```{r}
### Movie Bias
movie_bias <- colMeans(UI_train,na.rm = T) - mean_value
movie_bias %>% t %>% t  %>% kable(caption = "MOVIE BIAS")%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

### user bias
user_bias <- rowMeans(UI_train,na.rm = T) - mean_value
user_bias %>% t %>% t  %>% kable(caption = "USER BIAS")%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```


### From the raw average, and the appropriate user and item biases, calculate the baseline predictors for every user-item combination.

```{r}
### start from the matrix of the mean_rating
baseline_predictor <- mean_rating
minrating = 1
maxrating = 5
for (r in 1:nrow(baseline_predictor))
  for (c in 1:ncol(baseline_predictor))
    baseline_predictor[r,c] <- 
  ### We have to ensure that the results are in the range [minrating,maxrating]
  ### which is why we have the min(max()) wrapper
  min(
    max(
      baseline_predictor[r,c] + movie_bias[c] + user_bias[r],
      1),
    5)
     
baseline_predictor %>% kable()%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

```



## Calculate the RMSE for the baseline predictors for both your training data and your test data.

```{r}
### Training RMSE 
train_RMSE_baseline <- sqrt( mean ( (UI_train-baseline_predictor)^2, na.rm=T ) )
train_RMSE_baseline

### Test RMSE 
test_RMSE_baseline <- sqrt( mean ( (UI_test-baseline_predictor)^2, na.rm=T ) )
test_RMSE_baseline

```


## Summary results.
Lets calculate the percentage improvements based on the original (simple average) and baseline predictor (including bias) RMSE numbers for both Test and Train data sets.

```{r}
### improvement in TRAIN RMSE when moving from raw average to baseline predictor
train_RMSE_improvement = 1 - train_RMSE_baseline/ train_RMSE_raw
train_RMSE_improvement

### improvement in Test RMSE when moving from raw average to baseline predictor
test_RMSE_improvement  = 1 - test_RMSE_baseline / test_RMSE_raw
test_RMSE_improvement

```

The training RMSE declined from `r round(train_RMSE_raw,3)` to `r round(train_RMSE_baseline,3)`,
which is an improvement of `r round(100*train_RMSE_improvement,3)` percent.

The testing RMSE declined from `r round(test_RMSE_raw,3)` to `r round(test_RMSE_baseline,3)`,
which is an improvement of `r round(100*test_RMSE_improvement,3)` percent.






