---
title: "Data 612 Project 2 - Content-Based and Collaborative Filtering"
author: "Vinayak Patel"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    highlight: tango
    theme: united
  pdf_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(tinytex.verbose = TRUE)
```


```{r message=FALSE, warning=FALSE}
if (!require("knitr")) install.packages("knitr")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("kableExtra")) install.packages("kableExtra")
if (!require("dplyr")) install.packages("dplyr")
if (!require("Matrix")) install.packages("Matrix")
if (!require("recommenderlab")) install.packages("recommenderlab")
if (!require("gridExtra")) install.packages("gridExtra")
if (!require("graphics")) install.packages("graphics")
```

<p>&nbsp;</p>

## Introduction

**For assignment 2, start with an existing dataset of user-item ratings, such as our toy books dataset, MovieLens, Jester [http://eigentaste.berkeley.edu/dataset/] or another dataset of your choosing. Implement at least two of these recommendation algorithms:**

<span style="font-size:16px;"> 
• Content-Based Filtering
• User-User Collaborative Filtering
• Item-Item Collaborative Filtering
</span>

**As an example of implementing a Content-Based recommender, you could build item profiles for a subset of MovieLens movies from scraping http://www.imdb.com/ or using the API at https://www.omdbapi.com/ (which has very recently instituted a small monthly fee). A more challenging method would be to pull movie summaries or reviews and apply tf-idf and/or topic modeling.**

<span style="font-size:16px;"> You should evaluate and compare different approaches, using different algorithms, normalization techniques, similarity methods, neighborhood sizes, etc. You don’t need to be exhaustive—these are just some suggested possibilities.</span>


<span style="font-size:16px;"> You may use the course text’s recommenderlab or any other library that you want.</span>

## Data 

**Load your data into (for example)** 
<span style="font-size:16px;"> </span>
<span style="font-size:16px;"> I use MovieLens small datasets: 100,000 ratings and 3,600 tag applications applied to 9,000 movies by 600 users. </span>

```{r}
ratings_data <- read.csv('ratings.csv', stringsAsFactors = F)
movie_data <- read.csv('movies.csv', stringsAsFactors = F)
```


```{r echo=TRUE, paged.print=TRUE}
#Load results
head(ratings_data) %>% kable()%>%
  kable_styling(bootstrap_options = "striped", full_width = F)

head(movie_data)%>% kable()%>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```


```{r}
### Structure
str(ratings_data)
str(movie_data)
### Summary
str(movie_data)
summary(movie_data)
```

## Transform Data

I used `realRatingMatrix` from 'recommenderlab' to transform data.

```{r}

#### use pivot_wider to make a user matrix
ratings_data$userId <- as.factor(ratings_data$userId)

UI <- as(ratings_data, "realRatingMatrix")
dim(UI@data)
```

<span style="font-size:18px;"> First, I will look at the similarity between items. The more Yellow the cell is, the more similar two items are. Note that the diagonal is Red, since it’s comparing each items with itself. </span>

### Items

```{r}
#Items
similarity_items <- similarity(UI[,1:10], method = "cosine", which = "items")
# The more red the cell is, the more similar two users are. 
image(as.matrix(similarity_items), main="Items Similarity (cosine)")
```


### Users

<span style="font-size:18px;"> User based collaborative filtering algorithms are based on measuring the similarity between users.</span>
<span style="font-size:18px;"> First, I will look at the similarity between users. The more yellow the cell is, the more similar two users are.</span>


```{r}
#Users
similarity_users <- similarity(UI[1:10, ], method = "cosine", which = "userId")

# The more red the cell is, the more similar two users are. 
image(as.matrix(similarity_users), main="User Similarity (cosine)")
```

### Distribution of The Ratings

```{r}
#Distribution plot
ratings_dis <- as.vector(UI@data)
ratings_dis <- ratings_dis[ratings_dis != 0]
ratings_dis <- as.data.frame(table(ratings_dis))

ggplot(ratings_dis, aes(x = ratings_dis, y= Freq)) + 
  geom_bar(stat= "identity") +
  geom_text(aes(label=Freq), position=position_stack(vjust = 0.5), color="white", size=4)+
  labs(y = "No. of Ratings", x = "Ratings") + ggtitle("Distribution of the ratings")+
  theme(plot.title = element_text(hjust = 0.5))
```


Users who rated only a few movies their ratings might be biased, so I can remove that.
```{r}
movies <- UI[rowCounts(UI) > 50, colCounts(UI) > 
    50]
```

In general, any ratings matrix, especially movie ratings matrix, is bound to have some bias. Some users may give higher ratings than others.

To see the bias distribution of ratings, we will plot average rating per user.
```{r}
avg <- rowMeans(movies)
ggplot() + aes(avg) + geom_histogram(binwidth = 0.1) + xlab("Average Rating") + 
    ylab(" No. of Ratings")
```
We can see from the avg ratings distribution plot below that it varies a lot.

recommender lab normalizes the data when building a model. Let us normalize the ratings and confirm that all averages are 0 now to see what kind of effect it has.

```{r}
movie_Normalization <- normalize(movies)
avg <- round(rowMeans(movie_Normalization), 5)
table(avg)
```

```{r}
min_Items <- quantile(rowCounts(movies), 0.95)
min_Users <- quantile(colCounts(movies), 0.95)

image(movies[rowCounts(movies) > min_Items, colCounts(movies) > min_Users], 
    main = "Heatmap of the Top Users and Movies (Non-Normalized")
```
```{r}
image(movie_Normalization[rowCounts(movie_Normalization) > min_Items, colCounts(movie_Normalization) > 
    min_Users], main = "Heatmap of the Top Users and Movies (Normalized)")
```


## Split training and test datasets.

Split the dataset into training set (80%) and testing set (20%).

```{r}
set.seed(10)
train_set <- sample(x = c(TRUE, FALSE), size = nrow(movies), replace = TRUE, 
    prob = c(0.8, 0.2))

movie_Train <- movies[train_set, ]
movie_Test <- movies[!train_set, ]
```


##Item-Item Collaborative Filtering

I build an item-item collaborative filtering where I recommend movies to users where their item’s ratings are similar.

Using your training data, let’s create a model using method `IBCF` .

```{r}
IBCF_model <- Recommender(movie_Train, method = "IBCF")
```


```{r}
### Recommendations using test set
IBCF_pred <- predict(IBCF_model, newdata = movie_Test, n = 6)
```


Now, Let’s extract recommenders
Movie Ratings for first user

```{r}
user1 <- as.data.frame(movie_Test@data[1, movie_Test@data[1, ] > 0])
colnames(user1) <- c("Rating")
user1[c("movieId")] <- as.integer(rownames(user1))

data <- movie_data %>% inner_join(user1, by = "movieId") %>% select(Movie = "title", 
    Rating) %>% arrange(desc(Rating))
knitr::kable(data, format = "html") %>% kableExtra::kable_styling(bootstrap_options = c("striped", 
    "hover"))
```

### From the raw average, and the appropriate user and item biases, calculate the baseline predictors for every user-item combination.

```{r}
recommended <- IBCF_pred@itemLabels[IBCF_pred@items[[1]]]
recommended <- as.data.frame(as.integer(recommended))
colnames(recommended) <- c("movieId")
data <- recommended %>% inner_join(movie_data, by = "movieId") %>% select(Movie = "title")
knitr::kable(data, format = "html") %>% kableExtra::kable_styling(bootstrap_options = c("striped", 
    "hover"))
```

 
 
### User-User Collaborative Filtering

Finally, I build a user-user collaborative filtering where I recommend movies to users based on how similar they are with other users.

```{r}
UBCF_model <- Recommender(movie_Train, method = "UBCF")
```

####Recommendations using test set
```{r}
UBCF_pred <- predict(UBCF_model, newdata = movie_Test, n = 6)
```

Let us consider the first user and look at his/her recommendations. The first user gravitated towards more critically acclaimed dramas and these recommendations are among the best movies produced.

```{r}
# Recommendations for the first user
UBCF_recommended <- UBCF_pred@itemLabels[UBCF_pred@items[[1]]]
UBCF_recommended <- as.data.frame(as.integer(UBCF_recommended))
colnames(UBCF_recommended) <- c("movieId")

data <- UBCF_recommended %>% inner_join(movie_data, by = "movieId") %>% select(Movie = "title")
knitr::kable(data, format = "html") %>% kableExtra::kable_styling(bootstrap_options = c("striped", 
    "hover"))
```

#### Normalized 
```{r}
Normalized_model <- Recommender(movie_Train, method = "UBCF", parameter = list(normalize = NULL))
Normalized_pred <- predict(Normalized_model, newdata = movie_Test, n = 6)
recommended <- Normalized_pred@itemLabels[Normalized_pred@items[[1]]]
recommended <- as.data.frame(as.integer(recommended))
colnames(recommended) <- c("movieId")
data <- recommended %>% inner_join(movie_data, by = "movieId") %>% select(Movie = "title")
knitr::kable(data, format = "html") %>% kableExtra::kable_styling(bootstrap_options = c("striped", 
    "hover"))
```


## Summary results.

For both Item-Item and User-User collaborative filtering, the recommendations which the user got was almost similar. Movie ratings extracted for the first user consisted of action genre movies. Therefore it was interesting to see that the movies recommended to the same user also consisted of same genre movies.

```{r}

```

